Gas-Efficient Parameter Verification for Blockchain Smart Contracts Using Off-Chain Validators

### 2024-10-12 papers/assertion-contract/fc2025/paper.tex
submitted version

\lstdefinelanguage{solidity}{
  backgroundcolor=\color{pink!20},
  keywords={contract, function, public, int},
  keywordstyle=\bfseries\color{blue},
  commentstyle=\itshape\color{green!40!black},
  basicstyle=\ttfamily,
  morecomment=[l]{//},
  morecomment=[s]{/*}{*/},
  morestring=[b]",
  sensitive=true
}
\begin{document}
%
\title{Gas-Efficient Parameter Verification for Blockchain Smart Contracts Using Off-Chain Validators}
%
\titlerunning{Off-Chain Parameter Verification Approach: Gas-Efficient}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Thi Thu Ha Doan\orcidID{0000-0001-7524-4497}\and Peter Thiemann\orcidID{0000-0002-9000-1239}}
%
\authorrunning{Ha Doan, P. Thiemann}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{University of Freiburg, Germany \\
  \email{\{doanha,thiemann\}@informatik.uni-freiburg.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
In certain blockchain scenarios, verifying the properties of parameters may require significant gas costs or be infeasible due to gas limits. This study proposes a distributed verification protocol to address these challenges. The core concept involves offloading the verification process to validators, who attempt to find counterexamples or proofs off-chain. If a counterexample is discovered, it can be submitted for re-verification on-chain. If validated, the parameter is discarded; otherwise, the correctness of the parameter is guaranteed through a proof-of-work incentive mechanism, encouraging validators to participate in the distributed verification process. To implement these ideas, we propose a practical model that also tackles potential security concerns. We developed a prototype of the protocol and conducted a cost analysis, demonstrating its gas efficiency and practical utility.
\keywords{blockchain, off-chain verification, gas optimization, proof-of-work, distributed verification}
\end{abstract}
%
%
%
\section{Introduction}
\label{sec:introduction}
\lstset{language=solidity}
A smart contract \cite{szabo1997smartcontracts,eth-whitepaper,tezos-whitepaper} is a critical component of blockchain technology, with applications ranging from simple tasks like auctions to more complex systems such as decentralized markets. However, broader adoption of smart contracts faces challenges, with gas costs being among the most significant. Every computation performed by a smart contract on the blockchain incurs a cost, as each unit of computation and storage consumed by an algorithm must be paid for due to the burden it places on the system \cite{eth-whitepaper,eth-yellowpaper,gas-cost}.

@@

Consider a contract that takes a prime number as a parameter:
\begin{lstlisting}[numbers=none]
contract Example {
  function (int p) public {
    // assume p is prime 
  ... }
}
\end{lstlisting}
This assumption can be expressed as an explicit assertion in predicate logic:
\begin{gather*}
\label{eq1}
  (\forall n) (2 \leq n \leq \sqrt{p}) \Rightarrow (p \mathbin{\%} n) \ne 0
\end{gather*}
Checking the validity of this assumption requires a loop in the contract. The test would take \(O(\sqrt{p})\) time (assuming constant time for computing the remainder) and produce extra cost linear in \(\sqrt{p}\).

Similarly, consider a contract that requires a sorted array of integers as input:

\begin{lstlisting}[numbers=none]
contract Sorted {
  function find (int[100] a, int v) public {
    // assume a is sorted 
  ... } 
}
\end{lstlisting}
The explicit assertion would be:
\begin{gather*}\label{}
  (\forall k) (0\leq k <99) \Rightarrow a[k] \leq a[k+1]
\end{gather*}
This check can be performed in \(O(1)\) time, but the associated constant factor makes it costly.

@@

For the sorted array example, the negation is:
\begin{gather*}\label{}
  (\exists k) (0\leq k <99) \wedge a[k] > a[k+1]
\end{gather*}
Again, we can have every validator generate a random number \(k\). If the condition is true for such \(k\), the validator has found a counterexample for the sortedness of the array. Otherwise, the validator relies on others to check different indices.

@@

The paper is structured as follows: The next section defines the parameter verification problems and presents use cases. Section \ref{sec:offchain-distributed-verification-approach} proposes the novel approach, while Section \ref{sec:practical-model} introduces its practical model. Section \ref{sec:gas-cost-analysis} analyzes the gas costs, comparing our proposed approach with the on-chain method. Section \ref{sec:prototype-implementation} details the prototype implementation. Section \ref{sec:related-work} discusses related work, and Section \ref{sec:conclusion} concludes the paper.
\section{Parameter Assertion Verification On-chain}
\label{sec:parameter-assertion-verification-onchain}
Our work focuses on parameters that are computationally expensive to verify on-chain. Specifically, we address cases where verifying a parameter assertion requires loops within the contract and the assertion can be formalized using universal or existential quantifiers.
\subsection{Parameter Assertion Formalization}
\paragraph{Universal Quantifier Formalization}
An assertion can be formalized using the universal quantifier as follows:
\begin{gather}
  \label{eq:1}
  \forall a \in A. \, P(a)
\end{gather}
This formula can be extended to assertions over multiple domains as:
\begin{gather}
  \label{eq:2}
  \forall a \in A, \, \forall b \in B, \, \dots \, P(a, b, \dots)
\end{gather}
\paragraph{Existential Quantifier Formalization}
In cases where an assertion involves the existence of an element, it can be formalized using existential quantifiers as follows:
\begin{gather}
  \label{eq:3}
  \exists a \in A. \, P(a)
\end{gather}
This formula can also be extended to assertions over multiple domains:
\begin{gather}
  \label{eq:4}
  \exists a \in A, \, \exists b \in B, \, \dots \, P(a, b, \dots)
\end{gather}
\subsection{On-chain Verification}
Verifying these assertions on-chain requires iterating over the domain \(A\) within the contract. For the fomulas \ref{eq:5}, the verification fomulas is as follows.
\begin{gather}
  \label{eq:5}
  \bigwedge_{a \in A} P(a)
\end{gather}
For multiple domains, the verification process becomes:
\begin{gather}
  \label{eq:6}
  \bigwedge_{a \in A, b \in B, \dots} \dots P(a, b, \dots)
\end{gather}
Verifying these assertions in  the existential quantifier formalization requires iterating over the domain \(A\) until finding an element \(a\) that satisfies the predicate:
\begin{gather}
  \label{eq:7}
  \bigvee_{a \in A} P(a)
\end{gather}

For multiple domains, the verification process becomes:
\begin{gather}
  \label{eq:8}
  \bigvee_{a \in A, b \in B, \dots} \dots P(a, b, \dots)
\end{gather}
When the domain \(A\) is large or the predicate \(P\) is complex to evaluate, performing the verification on-chain incurs significant gas costs or even make the verification process infeasible due to gas limits. Given the current high gas prices, this can result in substantial costs.
\subsection{Use Cases}
Smart contract languages like Solidity \cite{solidity}, Vyper \cite{vyper}, Michelson \cite{michelson,michelson1}, and Plutus \cite{plutus} support common data types such as numbers, booleans, strings, bytes, arrays, sets, and maps. Many properties of these types can be expressed using universal quantifiers, requiring loops for verification in smart contracts, but can often be disproven by a counterexample or proven with a proof. Below are key properties for each type:
\paragraph{\textbf{Numbers}}
Number types exhibit some properties that are crucial in cryptographic applications, such as primality and coprimeness. For example, the condition for two numbers \(a\) and \(b\) to be coprime is:
\begin{gather*}
\label{}
(\forall n \in \mathbb{N}) \ (2 \leq n \leq \min(a, b)) \Rightarrow \neg((a \mod n = 0) \wedge (b \mod n = 0))
\end{gather*}
\paragraph{\textbf{Booleans}}
SAT problems involve determining if there exists an assignment of variables that makes a formula \( F \) in propositional logic true. Verifying if a given assignment satisfies a formula \( F \), which is usually in Conjunctive Normal Form (CNF), requires looping over variables and clauses. Although finding a satisfying assignment is NP-complete, verifying an assignment takes \( O(n \times m) \), where \( n \) is the number of clauses and \( m \) is the number of variables.
\paragraph{\textbf{Strings}}
String properties that require iteration include uniqueness (no repeated characters), character set membership, lexicographical order, pattern matching, and forbidden substrings. For instance, verifying that a string \(s\) does not contain a forbidden substring \(f\) of length \(k\) can be expressed as:
\begin{gather*}
(\forall i) \ (0 \leq i \leq n - k) \Rightarrow s[i, i+1, \dots, i+k-1] \neq f
\end{gather*}
\paragraph{\textbf{Bytes}}
Verifying byte arrays often requires loops for properties like uniformity, sorted order, specific byte patterns, uniqueness, valid range, and alternating patterns. For example, to check if all bytes in an array \(b\) lie within a range \([l, u]\), the condition is:
\begin{gather*}
(\forall i) \ (0 \leq i \leq n) \Rightarrow l \leq b[i] \leq u
\end{gather*}
\paragraph{\textbf{Arrays (or Lists)}}
Common properties of arrays include sortedness, uniqueness, boundedness, monotonicity, symmetry, cyclic structure, and constant values. %A tree-based structure like a max-heap, where each parent node is greater than or equal to its children, can be checked as:
Tree-based structures like heaps can also be built from arrays, with properties such as the heap property. For example, in a max-heap, each parent node is greater than or equal to its children. This property can be expressed as:
\begin{gather*}
(\forall i) \ a[i] \geq a[2i+1] \land a[i] \geq a[2i+2]
\end{gather*}
\paragraph{\textbf{Sets}}
Set properties like subset verification, disjoint sets, element uniqueness, set equality, and set intersection often require loops. For example, verifying that two sets \(A\) and \(B\) share at least one element can be expressed as:
\begin{gather*}
(\exists a \in A) (\exists b \in B)\ a = b
\end{gather*}
Such properties can be computationally expensive for large sets.
\section{Off-chain Distributed Verification Approach}
\label{sec:offchain-distributed-verification-approach}
\subsection{The Outline of the Approach}
We propose a novel approach for verifying parameter assertions in blockchain-based smart contracts. Our approach distributes the verification process off-chain, submitting only the outcomes to the blockchain for re-verification. This strategy aims to significantly reduce the computational burden on the blockchain while maintaining the integrity of the contract.

Our approach takes into account the key principle that assertions expressed with universal quantifiers can be disproven by counterexamples, while those with existential quantifiers can be proven with valid proofs. These counterexamples or proofs serve as checkpoints, representing specific elements within the domain.

The outline of our approach is as follows:
\paragraph{\textbf{Determine the checkpoint formula.}}
The first step is to establish the checkpoint formula used to identify counterexamples or proofs. For assertions with universal quantifiers, as shown in Equation~\ref{eq:1}, the checkpoint formula is the negation of the original assertion and is expressed using an existential quantifier as follows: 
\begin{gather}
\label{eq:9}
(\exists a \in A) \ \neg P(a)
\end{gather}
For assertions over multiple domains, as in Equation~\ref{eq:2}, the checkpoint formula becomes:
\begin{gather}
\label{eq:10}
(\exists a \in A) (\exists b \in B) \dots  \neg P(a, b, \dots)
\end{gather}
In the case of existential quantifier assertions, as in Equation~\ref{eq:3} and Equation~\ref{eq:4}, the checkpoint formula remains the same as the original assertion.
\paragraph{\textbf{Off-chain distributed effors.}}These checkpoint formulas are then distributed to validators. Off-chain, the validators attempt to find counterexamples or proofs by identifying some \( a \in A \) such that the checkpoint formula is satisfied. For universal quantifier formalizations, this means finding \( a \) where \( \neg P(a) \) holds, and for existential quantifier formalizations, finding \( a \) where \( P(a) \) is true.
\paragraph{\textbf{On-chain re-verification.}} Once a counterexample or proof involving an element \( a \) is found, the validator submits it to the blockchain, where it is re-verified by evaluating \( \neg P(a) \) or \( P(a) \) on-chain. For assertions with universal quantifiers, when a counterexample is successfully verified on-chain, the parameter assertion is proven to be invalid. Otherwise, it is considered valid, assuming that other points are checked by other validators. For assertions with existential quantifiers, the parameter assertion remains invalid until a proof is submitted and successfully re-checked on-chain.
\paragraph{\textbf{Incentivization Mechanism.}} Validators are incentivized to find counterexamples or provide proofs, which require computational resources. Since counterexamples are rare (as callers avoid submitting invalid parameters), confidence in correctness grows with more validator attempts at different checkpoints. Thus, incentivization is crucial for ensuring correctness in the absence of counterexamples.Validators are rewarded for both finding counterexamples and contributing computational proofs, with  counterexample discovery prioritized through higher rewards.

This approach reduces the blockchain's burden and lowers gas costs by verifying only specific checkpoints, rather than performing full on-chain verification.
\subsection{Proof-of-Work-Based Incentivization Mechanism}
The computational effort is represented by a computation proof, which is then submitted on-chain by the validator to claim their reward. This approach encourages participation in the distributed verification process, ensuring thorough validation even without the discovery of counterexamples.

However, significant challenges arise regarding (1) how to recheck the counterexample and the computation efforts on-chain at a reasonable gas cost, (2) how to ensure a significant number of checkpoints have been performed before a computation reward is issued, in order to guarantee an expected probability of correctness for the parameter, and (3) how to ensure that validators actually perform the checkpoint formula off-chain when no counterexample is found but a computation reward is provided.

To overcome these challenges, we propose a proof-of-work-based incentivization mechanism. We utilize the proof-of-work concept \cite{dwork1992pow,nakamoto2008bitcoin} by introducing a target hash \( t \), which sets the required difficulty level to earn the computation reward. The basic idea is that each verification process introduces a target hash. The validator attempts to perform the negation check and compute a hash that is less than or equal to the given target hash. The objective of the validator is to find either a counterexample or a computation that produces a hash smaller than or equal to the target hash \( t \).
\subsubsection{Introduce a seed} To recheck the counterexample or computation proof on-chain at a reasonable gas cost, the counterexample or computation proof should be submitted on-chain in the form of the indication of the element \( a \in A \) where the assertion is invalid, in the case of a counterexample. The on-chain re-verification process is then straightforward, simply requiring the evaluation of \( \neg P(a) \).

The next challenge is how to generate a random hash for each computation and devise a computational proof to incentivize computational effort. To address this, we introduce a seed \( s \) in each verification process.

Let \( n \) denote the size of the domain \( A \). 
\begin{gather*}
n = |A|
\end{gather*}
where \(|A| \) denotes the number of elements in the domain $A$.

Each number \( i \) in the range \([1, n]\) corresponds to an element in \( A \). 
\begin{gather*}
a=A[i]
\end{gather*}
where $A[i]$ denotes the $i$-th element in $A$.

The seed \( s \) is a random number. 
\begin{gather*}
s \in \mathbb{N} 
\end{gather*}
The seed $s$  is then reduced with modulo \( n \) to obtain \( i \), which is associated with an element \( a \) in \( A \).
\begin{gather}
\label{eq:11}
i = (s\mod n) + 1
\end{gather}
\subsubsection{Estimate the number of seeds:} To ensure that a sufficient number of checkpoints have been checked before a computational proof is found, we first estimate the number of seeds required to find a counterexample with the desired probability of correctness \cite{ross2014probability,feller1968probability}.

@@

To determine the number of seeds needed based on the expected probability of detecting a problem, we define \( p \) as the expected probability of finding at least one counterexample. 
\begin{itemize}
    \item The probability of a single seed failing to find a counterexample is \( \frac{n-1}{n} \).
    \item The probability of all \( m \) seeds failing is \( \left( \frac{n-1}{n} \right)^m \).
    \item Thus, the probability of at least one seed finding a counterexample is:
\begin{gather*}
  p = 1 - \left( \frac{n-1}{n} \right)^m
\end{gather*}
\end{itemize}
%Rearranging this gives:
%\begin{gather*}
%\left( \frac{n-1}{n} \right)^m = 1 - p
%\end{gather*}
%Taking the natural logarithm:
%\begin{gather*}
%m \ln\left( \frac{n-1}{n} \right) = \ln(1 - p)
%\end{gather*}
Solving for \( m \):
\begin{gather}
\label{eq:12}
m = \frac{\ln(1 - p)}{\ln\left( \frac{n-1}{n} \right)}
\end{gather}

So the number of the seeds to achive the expected propety $p$ is great or at least equal to $m$. Note that, \( p \) represents the probability when only one validator participates. As more validators are involved, the probability of discovering a counterexample increases.
\subsubsection{Set up the taget hash:}
The above number of seed $m$ can then be used to set up the target hash \cite{nakamoto2008bitcoin,bonneau2015sok}. The target hash $t$ can be calculated based on a desired difficulty level. In proof-of-work systems, difficulty is typically adjusted to ensure that the average efforts to find a valid hash. The target hash $t$ is often calculated by taking the maximum possible hash value and dividing it by the difficulty level.

Given the expected number of seeds to generate, 
the difficulty level \( D \) can be adjusted accordingly. Typically, \( D \) is set to ensure that a valid hash is found approximately once every \( m \) trials.

If we want the probality of finding a valid hash to be $\frac{1}{m}$, then \( D \) equal to \( m \):
\begin{gather}
\label{eq:13}
t = \frac{H_{\text{max}}}{m}
\end{gather}

Where \( H_{\text{max}} \) is the maximum possible hash value and \( m \) is the difficulty level, which can be adjusted based on the expected number of trials.
\subsubsection{Conduct a Computation Hash:} To ensure that a computational proof includes the negation check, the computation of the checkpoint formula is integrated into the proof process. This guarantees that validators must execute the negation check as part of producing the computation proof.

For each element \( a \), let \(\pi_{P_{a}}\) be the function that returns the essential components of evaluating the negation \(\neg P(a)\). The computation hash thus incorporates both the seed \( s \) and \(\pi_{P_{a}}\), ensuring proper validation.

\begin{gather}
\label{eq:14}
t_s = \text{hash}(s, \pi_{P_{a}})
\end{gather}
where \(\text{hash}\) is a cryptographic hash function used to compute the hash value.

Assertions in smart contract languages, such as Solidity and Michelson, are often represented as predicate logic, expressed through Boolean formulas. These are composed of relational operators like \( =, \neq, >, <, \geq, \leq \), logical operators such as \( \neg, \land, \lor\), and corresponding expressions.

Thus, \(\pi_{P_{a}}\) represents the set of expressions involved in evaluating the predicate \(P_a\).
\begin{gather*}
\pi_{P_{a}} = \{ e_1, e_2, \ldots \}
\end{gather*}
where \(e_i\) are the expressions that contribute to the computation of the predicate \(P_a\).

%For example, for each element \( i \), it returns \(a[i]\) and \(a[i + 1]\) in the evaluation of the negation predicate \( a[i] < a[i+1] \) in the sorted array example.

>>>>

\lstdefinelanguage{solidity}{
  backgroundcolor=\color{pink!20},
  keywords={contract, function, public, int},
  keywordstyle=\bfseries\color{blue},
  commentstyle=\itshape\color{green!40!black},
  basicstyle=\ttfamily,
  morecomment=[l]{//},
  morecomment=[s]{/*}{*/},
  morestring=[b]",
  sensitive=true
}
\begin{document}
%
\title{Gas-Efficient Parameter Verification for Blockchain Smart Contracts Using Off-Chain Validators}
%
\titlerunning{Off-Chain Parameter Verification Approach: Gas-Efficient}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
%\author{Thi Thu Ha Doan\orcidID{0000-0001-7524-4497}\and Peter Thiemann\orcidID{0000-0002-9000-1239}}
%
%\authorrunning{Ha Doan, P. Thiemann}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
%\institute{University of Freiburg, Germany \\
%  \email{\{doanha,thiemann\}@informatik.uni-freiburg.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
In certain blockchain scenarios, verifying the properties of parameters may require significant gas costs or be infeasible due to gas limits. This study proposes a distributed verification protocol to address these challenges. The core concept involves offloading the verification process to validators, who attempt to find counterexamples or proofs off-chain. If a counterexample is discovered, it can be submitted for re-verification on-chain. If validated, the parameter is discarded; otherwise, the correctness of the parameter is guaranteed through a proof-of-work incentive mechanism, encouraging validators to participate in the distributed verification process. To implement these ideas, we propose a practical model that tackles potential security concerns. We developed a prototype of the protocol and conducted a cost analysis, demonstrating its gas efficiency and practical utility.
\keywords{blockchain, off-chain verification, gas optimization, proof-of-work, distributed verification}
\end{abstract}
%
%
%
\section{Introduction}
\label{sec:introduction}
\lstset{language=solidity}
A smart contract \cite{szabo1997smartcontracts,eth-whitepaper,tezos-whitepaper} is a critical component of blockchain technology, with applications ranging from simple tasks like auctions to more complex systems such as decentralized markets. However, broader adoption of smart contracts faces challenges, with gas costs being among the most significant. Every computation performed by a smart contract on the blockchain incurs a cost, as each unit of computation and storage consumed by an algorithm must be paid for due to the burden it places on the system \cite{eth-whitepaper,eth-yellowpaper,gas-cost}.

@@

Consider a contract that takes a prime number as a parameter:
\begin{lstlisting}[numbers=none]
contract Example {
  function (int p) public { 
    // assume p is prime 
  ... }}
\end{lstlisting}
This assumption can be expressed as an explicit assertion in predicate logic:
\begin{gather*}
\label{eq1}
  (\forall n) (2 \leq n \leq \sqrt{p}) \Rightarrow (p \mathbin{\%} n) \ne 0
\end{gather*}
Checking the validity of this assumption requires a loop in the contract. The test would take \(O(\sqrt{p})\) time (assuming constant time for computing the remainder) and produce extra cost linear in \(\sqrt{p}\).

Similarly, consider a contract that requires a sorted array \(a\)  of the size \(n\)  of integers as input. 
%\begin{lstlisting}[numbers=none]
%contract Sorted {
%  function find (int[100] a, int v) public {
%    // assume a is sorted
%  ...}}
%\end{lstlisting}
The explicit assertion would be:
\begin{gather*}\label{}
  (\forall k) (0\leq k <n) \Rightarrow a[k] \leq a[k+1]
\end{gather*}
This check can be performed in \(O(1)\) time, but the associated constant factor makes it costly.

@@

For the sorted array example, the negation is:
\begin{gather*}\label{}
  (\exists k) (0\leq k <n) \wedge a[k] > a[k+1]
\end{gather*}
Again, we can have every validator generate a random number \(k\). If the condition is true for such \(k\), the validator has found a counterexample for the sortedness of the array. Otherwise, the validator relies on others to check different indices.

@@

The paper is structured as follows: The next section defines the parameter verification problems and presents use cases. Section \ref{sec:offchain-distributed-verification-approach} proposes the novel approach, while Section \ref{sec:practical-model} introduces its practical model. Section \ref{sec:gas-cost-analysis} analyzes the gas costs, comparing our proposed approach with the on-chain method. Section \ref{sec:prototype-implementation} details the prototype implementation. Section \ref{sec:related-work} discusses related work, and Section \ref{sec:conclusion} concludes the paper.
\section{Parameter Assertion Verification On-chain}
\label{sec:parameter-assertion-verification-onchain}
Our work targets parameters that are computationally expensive to verify on-chain, particularly when verification involves loops within the contract. We focus on assertions that can be expressed using universal or existential quantifiers.
\subsection{Parameter Assertion Formalization}
\paragraph{Universal Quantifier Formalization}
An parameter assertion can be formalized using the universal quantifier over the domain \(A\) as:
\begin{equation} \label{eq:1}
    \forall a \in A, \, P(a)
\end{equation}
For multiple domains:
\begin{equation} \label{eq:2}
    \forall a \in A, \, \forall b \in B, \, \dots, \, P(a, b, \dots)
\end{equation}

\paragraph{Existential Quantifier Formalization}
For assertions involving the existence of an element, we use existential quantifiers:
\begin{equation} \label{eq:3}
    \exists a \in A, \, P(a)
\end{equation}
Extended to multiple domains:
\begin{equation} \label{eq:4}
    \exists a \in A, \, \exists b \in B, \, \dots, \, P(a, b, \dots)
\end{equation}
\subsection{On-chain Verification}
Verifying these assertions on-chain involves iterating over the domain \(A\)\footnote{For multiple domains, the verification process requires iterating over all domains} within the contract. For formula \ref{eq:1}, the verification is expressed as: \(   \bigwedge_{a \in A} P(a)\).

´For the existential quantifier formalization as in formula \ref{eq:3}, verification requires iterating over \(A\) until an element \(a\) is found that satisfies \(P\): \(\bigvee_{a \in A} P(a)\).

When \(A\) is large or \(P\) is complex to evaluate, this process can incur significant gas costs or even become infeasible due to gas limits. Given the current high gas prices, this could result in substantial expenses.
\subsection{Use Cases}
Smart contract languages like Solidity \cite{solidity}, Vyper \cite{vyper}, Michelson \cite{michelson,michelson1}, and Plutus \cite{plutus} support common data types such as numbers, booleans, strings, bytes, arrays, sets. Many properties of these types use quantifiers, requiring loops for verification but can often be disproven by a counterexample or proven formally. Key properties for each type are outlined below:
\paragraph{\textbf{Numbers}}
Number types have some properties that are vital in cryptographic applications, such as primality and coprimeness. %For instance, the condition for two numbers \(a\) and \(b\) to be coprime is:
%\begin{equation*}
%(\forall n) \ (2 \leq n \leq \min(a, b)) \Rightarrow \neg((a \mod n = 0) \wedge (b \mod n = 0))
%\end{equation*}
\paragraph{\textbf{Booleans}}
SAT problems determine if there is an assignment of variables that makes a formula \( F \) in propositional logic true. Verifying whether a given assignment satisfies \( F \), typically in Conjunctive Normal Form (CNF), involves iterating over variables and clauses. While finding a satisfying assignment is NP-complete, verifying one takes \( O(n \times m) \), where \( n \) is the number of clauses and \( m \) is the number of variables.
\paragraph{\textbf{Strings}}
String properties that require iteration include uniqueness, character set membership, lexicographical order, pattern matching, and forbidden substrings. %For example, verifying that a string \(s\) lacks a forbidden substring \(f\) of length \(k\) is expressed as:
%\begin{gather*}
%(\forall i) \ (0 \leq i \leq n - k) \Rightarrow s[i, i+1, \dots, i+k-1] \neq f
%\end{gather*}
\paragraph{\textbf{Bytes}}
Verifying byte arrays often involves loops for properties such as uniformity, sorted order, specific byte patterns, uniqueness, valid range, and alternating patterns. For instance, to check if all bytes in an array \(b\) are within a range \([l, u]\), the condition is:
\begin{gather*}
(\forall i) \ (0 \leq i \leq n) \Rightarrow l \leq b[i] \leq u
\end{gather*}
\paragraph{\textbf{Arrays (or Lists)}}
Common properties of arrays include sortedness, uniqueness, boundedness, monotonicity, symmetry, cyclic structure, and constant values. %A tree-based structure like a max-heap, where each parent node is greater than or equal to its children, can be checked as:
Tree-based structures like heaps can also be built from arrays, with properties such as the heap property. For example, in a max-heap, each parent node is greater than or equal to its children. This property can be expressed as:
\begin{gather*}
(\forall i) \ a[i] \geq a[2i+1] \land a[i] \geq a[2i+2]
\end{gather*}
\paragraph{\textbf{Sets}}
Set properties like subset verification, disjoint sets, element uniqueness, set equality, and set intersection often require loops. For example, verifying that two sets \(A\) and \(B\) share at least one element can be expressed as:
\begin{gather*}
(\exists a \in A) (\exists b \in B)\ a = b
\end{gather*}
%Such properties can be computationally expensive for large sets.
\section{Gas-Efficient Off-chain Distributed Parameter Verification Approach}
\label{sec:offchain-distributed-verification-approach}
\subsection{The Outline of the Approach}
We propose a novel approach for verifying parameter assertions in blockchain-based smart contracts by offloading the verification process off-chain and submitting only the results for on-chain re-verification. 
Our approach leverages the principle that assertions with universal quantifiers can be disproven by counterexamples, while those with existential quantifiers can be proven with valid proofs. These counterexamples or proofs act as checkpoints, representing specific elements within the domain.

The outline of our approach is as follows:
\paragraph{\textbf{Determine the checkpoint formula.}}
The first step is to establish the checkpoint formula used to identify counterexamples or proofs. For assertions with universal quantifiers, the checkpoint formula is the negation of the original assertion and is expressed using an existential quantifier as follows: 
\begin{gather}
\label{eq:9}
(\exists a \in A) \ \neg P(a)
\end{gather}
For assertions over multiple domains, the checkpoint formula becomes:
\begin{gather}
\label{eq:10}
(\exists a \in A) (\exists b \in B) \dots  \neg P(a, b, \dots)
\end{gather}
In the case of existential quantifier assertions, the checkpoint formula remains the same as the original assertion.
\paragraph{\textbf{Off-chain distributed effors.}} These checkpoint formulas are distributed to validators, who then work off-chain to find counterexamples or proofs by identifying some \( a \in A \) that satisfies the checkpoint formula. For universal quantifiers, this involves finding \( a \) where \( \neg P(a) \) holds, while for existential quantifiers, it requires finding \( a \) where \( P(a) \) is true.
\paragraph{\textbf{On-chain re-verification.}} After finding a counterexample or proof involving an element \( a \), the validator submits it to the blockchain for on-chain re-verification by evaluating \( \neg P(a) \) or \( P(a) \). For universal quantifiers, a verified counterexample invalidates the assertion; otherwise, it is assumed valid if other points are verified by other validators. For existential quantifiers, the assertion remains invalid until a proof is submitted and verified on-chain.
\paragraph{\textbf{Incentivization Mechanism.}} Validators are incentivized to find counterexamples or provide proofs, which require computational resources. Since counterexamples are rare (as callers avoid submitting invalid parameters), confidence in correctness grows with more validator attempts at different checkpoints. Thus, incentivization is crucial for ensuring correctness in the absence of counterexamples.Validators are rewarded for both finding counterexamples and contributing computational proofs, with  counterexample discovery prioritized through higher rewards.

Overall, the approach reduces the blockchain's computational load while preserving contract integrity. By verifying only specific checkpoints instead of full on-chain verification, it lowers the blockchain's burden and reduces gas costs.
\subsection{Proof-of-Work-Based Incentivization Mechanism}
The computational effort is represented by a computation proof, submitted on-chain by the validator to claim a reward. This incentivizes participation in distributed verification, ensuring thorough validation even without finding counterexamples.

Challenges include: (1) rechecking counterexamples and computation efforts on-chain at low gas costs, (2) ensuring sufficient checkpoints are performed before rewarding computation, guaranteeing a probability of correctness, and (3) verifying that validators perform the checkpoint formula off-chain when no counterexample is found.

To address these, we propose a proof-of-work-based mechanism. Using the proof-of-work concept \cite{dwork1992pow,nakamoto2008bitcoin}, a target hash \( t \) sets the difficulty level for earning rewards. Each verification introduces a target hash, and the validator performs the checkpoint, computing a hash. The validator’s goal is to find a counterexample or produce a hash  that is less than or equal to \( t \).
\subsubsection{Introduce a seed} 
Let the element \( a \in A \) be where the assertion fails, either as a counterexample or as the computation proof found. To recheck this on-chain at a reasonable gas cost, it should be submitted as the element \( a \in A \) or as an indication of it. The on-chain re-verification then simply evaluates \( \neg P(a) \). The next challenge is generating a random hash for each checkpoint computation to provide computational proof. We address these by introducing a seed \( s \) in each verification process.

Let \( n \) denote the size of the domain \( A \), where \( n = |A| \) and \(|A|\) is the number of elements in \( A \). Each number \( i \) in the range \([1, n]\) corresponds to an element in \( A \), such that \( a = A[i] \), where \( A[i] \) denotes the \( i \)-th element of \( A \).

We introduce a seed \( s \), a random number \( s \in \mathbb{N} \). The seed \( s \) is reduced modulo \( n \) to obtain \( i \), which is then associated with an element \( a \) in \( A \):
\begin{equation}
\label{eq:11}
i = (s \mod n) + 1
\end{equation}
\subsubsection{Estimate the Number of Seeds:} To ensure a sufficient number of checkpoints are checked before finding a computational proof, we estimate the number of seeds required to find a counterexample with the desired probability of correctness \cite{ross2014probability,feller1968probability}.

@@

To determine the number of seeds needed based on the expected probability of detecting a problem, we define \( p \) as the expected probability of finding at least one counterexample. 
\begin{itemize}
    \item The probability of a single seed failing to find a counterexample is \( \frac{n-1}{n} \).
    \item The probability of all \( m \) seeds failing is \( \left( \frac{n-1}{n} \right)^m \).
    \item Thus, the probability of at least one seed finding a counterexample is:
\begin{gather*}
  p = 1 - \left( \frac{n-1}{n} \right)^m
\end{gather*}
\end{itemize}
Solving for \( m \):
\begin{equation}
\label{eq:12}
m = \frac{\ln(1 - p)}{\ln\left( \frac{n-1}{n} \right)}
\end{equation}

Therefore, the number of seeds required to achieve the expected probability \( p \) is at least \( m \). Note that \( p \) represents the probability when only one validator participates. As more validators are involved, the probability of discovering a counterexample increases.
\subsubsection{Set up the Target Hash:}
The number of seeds \( m \) can be used to set up the target hash \cite{nakamoto2008bitcoin,bonneau2015sok}. The target hash \( t \) is calculated based on a desired difficulty level. In proof-of-work systems, difficulty is typically adjusted to ensure that the average effort to find a valid hash matches the desired rate. The target hash \( t \) is often determined by taking the maximum possible hash value and dividing it by the difficulty level.

Given the expected number of seeds, the difficulty level \( D \) can be adjusted accordingly. Typically, \( D \) is set to ensure that a valid hash is found approximately once every \( m \) trials. If we want the probability of finding a valid hash to be \( \frac{1}{m} \), then \( D \) is set to \( m \):
\begin{equation}
\label{eq:13}
t = \frac{H_{\text{max}}}{m}
\end{equation}
Where \( H_{\text{max}} \) is the maximum possible hash value, and \( m \) is the difficulty level, adjusted based on the expected number of trials.
\subsubsection{Conduct a Computation Hash:} To ensure a computational proof includes the negation check, the checkpoint formula is integrated into the proof process, requiring validators to execute it as part of producing the proof.

For each element \( a \), let \(\pi_{P_{a}}\) represent the components for evaluating the checkpoint \(\neg P(a)\). The computation hash combines the seed \( s \) and \(\pi_{P_{a}}\) to ensure validation.

\begin{equation}
\label{eq:14}
t_s = \text{hash}(s, \pi_{P_{a}})
\end{equation}
where \(\text{hash}\) is a cryptographic hash function.

Assertions in smart contract languages are represented as predicate logic expressed through Boolean formulas. These formulas are composed of relational operators like \( =, \neq, >, <, \geq, \leq \) and logical operators such as \( \neg, \land, \lor \), along with corresponding expressions.

Thus, \(\pi_{P_{a}}\) denotes the set of expressions used to evaluate \(P_a\):
\begin{equation*}
\pi_{P_{a}} = \{ e_1, e_2, \ldots \}
\end{equation*}
where \(e_i\) are the expressions contributing to the computation of \(P_a\).
\subsection{Off-chain Distributed Verification}
For assertions using an existential quantifier, validation is straightforward: validators find a proof off-chain and submit it on-chain for re-verification. If valid, the parameter is confirmed, and the validator is rewarded.

For universal quantifier assertions, the process is more complex. The size \( n \) of the domain \( A \) is calculated, and the target hash is computed based on the expected probability of correctness. Validators generate approximately \( m \) seeds to find either a counterexample or a computational proof, confirming or invalidating the parameter.

As shown in Fig. \ref{fig:find_seed}, validators generate a seed \( s \) and perform the checkpoint until they find a counterexample or computation proof. They then submit the result on-chain to claim their reward.
